Discrete Quantum Theories and Computing

### 2018-01-25 prop_prop_letter_ajhs_referee_response.tex

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys, longbibliography, amsmath, amssymb, floatfix]{revtex4-1} 
\pdfoutput=1

\usepackage{fullpage}
\usepackage{cmap}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[french,main=english]{babel}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage[backref=false]{hyperref}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[braket,qm]{qcircuit}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\usepackage{comment}
\usepackage{xparse}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem{lemma}{Lemma}[thm]
\newtheorem{cor}{Corollary}[thm]
\newtheorem{assumption}{Assumption}
\newtheorem{fact}{Fact}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{condition}[assumption]{Condition}

\DeclareUnicodeCharacter{0229}{\c{e}}

\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\imposs}{{\text{\wesa{impossible}}}}
\newcommand{\likely}{{\text{\wesa{likely}}}}
\newcommand{\unlikely}{{\text{\wesa{unlikely}}}}
\newcommand{\necess}{{\text{\wesa{certain}}}}
\newcommand{\unknown}{{\text{\wesa{unknown}}}}
\newcommand{\midd}{{\text{\wesa{middle}}}}
\newcommand{\fket}[1]{{|#1\rangle}}
\newcommand{\fproj}[1]{|#1\rangle\langle #1|}
\newcommand{\proj}[1]{\op{#1}{#1}}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{L{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{R{#1}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\rmd}{d}
\newcommand{\rmi}{i}
\lineskiplimit=1pt
\newcommand{\ultramodular}{\mathcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\muB}{\ensuremath{\mu^{B}}}
\newcommand{\eventsC}{\ensuremath{\events_{C}}}

\newcommand{\says}[3]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{#1}{#2 says: #3}\end{minipage}\end{framed}}
\newcommand{\amr}[1]{\says{green}{Amr}{#1}}
\newcommand{\yutsung}[1]{\says{purple}{Yu-Tsung}{#1}}
\newcommand{\gerardo}[1]{\says{OliveGreen}{Gerardo}{#1}}
\newcommand{\andy}[1]{\says{blue}{Andy}{#1}}
% \excludecomment{suggest}
\NewDocumentEnvironment{suggest}{m m O{}}{#1{TEXT \##2 would go here. #3}}{#1{TEXT \##2 end here.}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Response of the Referee Report of Quantum Interval-Valued Probability:
Contextuality and the Born Rule}

\author{Yu-Tsung Tai}
\affiliation{Department of Mathematics, Indiana University, Bloomington, Indiana 47405,
USA}
\affiliation{Department of Computer Science, Indiana University, Bloomington,
Indiana 47405, USA}

\author{Andrew J. Hanson}
\affiliation{Department of Informatics, Indiana University, Bloomington,
Indiana 47405, USA}

\author{Gerardo Ortiz}
\affiliation{Department of Physics, Indiana University, Bloomington, Indiana 47405,
USA}

\author{Amr Sabry}
\affiliation{Department of Computer Science, Indiana University, Bloomington,
Indiana 47405, USA}

\date{\today}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Classical $\frac{1}{4}$-deterministic Probability Measure}

The first step in our formalization is to introduce a family of QIVPMs
parameterized by an uncertainty~$\delta$, which we call
$\delta$-deterministic QIVPMs.

\begin{definition}[$\delta$-Determinism]\label{def:delta-deterministic} A
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is
  $\delta$-deterministic if, for every event $P\in\events$, we have
  that either 
  $\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]$ or
  $\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]$. 
\end{definition}

\begin{table}
  \noindent \centering{}\caption{\label{tab:classical-IVPMs}Example
    QIVPMs~$\bar{\mu}_{A}$, $\bar{\mu}_{B}$ and $\bar{\mu}_{C}$ on
    commuting events $P$.}
\begin{tabular}{cccc}
\toprule
\addlinespace
$P$ & $\bar{\mu}_{A}(P)$ & $\bar{\mu}_{B}(P)$ & $\bar{\mu}_{C}(P)$\tabularnewline\addlinespace
\midrule
\midrule
\addlinespace
$\mathbb{0}$ & $\imposs$ & $\imposs$ & $\imposs$\tabularnewline\addlinespace
\midrule
\addlinespace
$\proj{0}$ & $\necess$ & $\left[\frac{2}{3},1\right]$ & $\left[\frac{2}{3},\frac{2}{3}\right]$\tabularnewline\addlinespace
\midrule
\addlinespace
$\proj{1}$ & $\imposs$ & $\left[0,\frac{1}{3}\right]$ & $\left[\frac{1}{3},\frac{1}{3}\right]$\tabularnewline\addlinespace
\midrule
\addlinespace
$\mathbb{1}$ & $\necess$ & $\necess$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\yutsung{$\bar{\mu}_{A}$ is neither a QIVPM nor a restriction of a QIVPM
according to the current definition. Need to fix this example or fix the definition.}
\end{table}

\noindent This definition puts no restrictions on the set of intervals
itself, only on which intervals are assigned to events. When
$\delta=0$, every event must be assigned a probability either in
$\imposs$ or in $\necess$, i.e., every event is completely determined
with certainty. As $\delta$ gets larger, the QIVPM allows for more
indeterminate behavior. 
For example, $\bar{\mu}_{A}$ defined in
Table~\ref{tab:classical-IVPMs} is $0$-deterministic because
$\imposs\subseteq\left[0,0\right]$ and
$\necess\subseteq\left[1-0,1\right]$.  By interpreting the states
~$\proj{0}$ and $\proj{1}$ as the head and the tail of a coin,
$\bar{\mu}_{A}$ represents tossing a double-headed coin. When $\delta$
is increased to $\frac{1}{3}$, both $\bar{\mu}_{B}$ and
$\bar{\mu}_{C}$ become $\frac{1}{3}$-deterministic. In contrast,
$\bar{\mu}_{A}$ is still not $\frac{1}{3}$-deterministic because
$\bar{\mu}_{A}(\proj{0})=\bar{\mu}_{A}(\proj{1})=\unknown=[0,1]$, but
both $\unknown\subseteq\left[0,\frac{1}{3}\right]$ and
$\unknown\subseteq\left[1-\frac{1}{3},1\right]$ are false.  Finally,
when $\delta=1$, because every interval is a subset of $\unknown$,
every QIVPM is $1$-deterministic. In particular, the non-informative
QIVPM~$\bar{\mu}_{A}$ is 1-deterministic.
\begin{comment}
and the Born QIVPM~$\bar{\mu}_{3}$ defined in
Eq.~(\ref{eq:Born-QIVPM}) are both $1$-deterministic.
\end{comment}

\paragraph{Why a Quantum $\frac{1}{4}$-deterministic Probability Measure is
Impossible}

Our framework allows us to generalize the above theorem to state that
for small enough $\delta$, it is impossible to have
$\delta$-deterministic QIVPMs. We next prove a main result of this
paper which is a stronger statement of contextuality that includes the
effects of finite-precision.  Every QIVPM must map some events to
truly uncertain intervals, not just ``almost definite intervals.'' 
\begin{comment}
\andy{contextuality is never defined.}
\yutsung{Actually, the second sentence in the introduction defines
``contextuality''. The problem is that there is no direct relation between what
defined in the introduction and here.}
\end{comment}

\begin{thm}[Finite-precision Extension of the Kochen-Specker
  Theorem] \label{cor:Kochen-Specker-IVPM} Given a Hilbert space
  $\Hilb$ of dimension~$d\ge3$, there is no $\delta$-deterministic
  QIVPM for $\delta<\frac{1}{3}$.\end{thm}

The proof is by contradiction: Suppose there is a
$\delta$-deterministic
QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$. Now use this
assumed QIVPM to construct the following 0-deterministic QIVPM
$\bar{\mu}^{\textrm{D}}:\events\rightarrow\left\{
  \imposs,\necess\right\}$:
\[
\bar{\mu}^{\textrm{D}}\left(P\right)=\begin{cases}
\imposs\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]\:;\\
\necess\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]\:.
\end{cases}
\]
However, since we know that 0-deterministic QIVPMs do not exist, the
map $\bar{\mu}^{\textrm{D}}$ cannot exist and hence $\bar{\mu}$ does
not exist. It remains to verify that $\bar{\mu}^{\textrm{D}}$ is
indeed a 0-deterministic QIVPM by checking that for orthogonal
projectors~$P_{0}$ and $P_{1}$, we have that
$\bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs$
implies
$\bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs$~\cite{TaiThesis2018}. Indeed
when
$\bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs$,
we have both $\bar{\mu}\left(P_{0}\right)$ and
$\bar{\mu}\left(P_{1}\right)\subseteq\left[0,\delta\right]$ which
implies
$\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\left[0,2\delta\right]$.
When $\delta<\frac{1}{3}$, i.e., $2\delta<1-\delta$, it follows that
$\bar{\mu}\left(P_{0}+P_{1}\right)$ cannot be a subset of
$\left[1-\delta,1\right]$. Since $\bar{\mu}\left(P_{0}+P_{1}\right)$
is a subset of either $\left[0,\delta\right]$ or
$\left[1-\delta,1\right]$ and the latter is excluded, then
$\bar{\mu}\left(P_{0}+P_{1}\right)$ must be a subset of
$\left[0,\delta\right]$, which implies
$\bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs$. Thus the
theorem holds when $\delta<\frac{1}{3}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{prop}
\end{document}

### 2018-02-06 dissertation.tex
Updates from ShareLaTeX
dissertation.tex Updating informations...

\documentclass[twoside]{iuphd}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[ backend=biber]{biblatex}
\addbibresource{prop.bib}
\ExecuteBibliographyOptions{sorting=none,maxbibnames=10,doi=false,isbn=false,url=false}
\DeclareUnicodeCharacter{0229}{\c{e}}
\DeclareUnicodeCharacter{00A0}{~}
\usepackage{mmap}
\usepackage{comment}

\begin{document}
%data for Title Page
\begin{comment}
\title{Dissertation Title}
\author{Author}
\date{Completion Date}
\end{comment}
\title{Discrete Quantum Theories and Computing}
\author{Yu-Tsung Tai}
\date{\today}
\maketitle

%data for Acceptance Page
\committeechair{Committee Chair}
\readertwo{2nd Reader}
\readerthree{3rd Reader}
\readerfour{4th Reader}
\defensedate{Defense Date}
\acceptancepage

%data for Copyright Page
\begin{comment}
\cryear{Current Year}
\end{comment}
\cryear{2018}
\copyrightpage

@@

\begin{abstract}
\begin{comment}
 The abstract is double-spaced and limited to 350 words. As many people will learn about
your work through your abstract published in Dissertation Abstracts (http://proquest.umi.com/login),
you should spend a good bit of effort in the composition of both the abstract and the title of your work.
Try to convey the flavor of your work, not just the bare bones of your findings; in an average abstract
there will be about 70 characters per line with a maximum of 35 lines. You should also work to phrase your
title so that it truly describes the contents and will be easily found in the index of Dissertation Abstracts.
The index is based on key words, so be as specific as you can be about your subject.
\end{comment}
Our primary research interest is to build a quantum computing model
characterizing realistic quantum computers. While most of the quantum
computing models based on uncomputable numbers, that is, the continuum
of real numbers, most of the classical computers in our daily life
are digital instead of analog computers. This highlight the necessity
to investigate discrete models for quantum theory and computing. Specifically,
we start from replacing the continuum of complex numbers by the discrete
finite fields. Although we have fruitful results on their geometric
implications and computing powers, their probability models are still
not completely satisfactory. To address this issue, we further exploited
quantum interval-valued probability, and proved an imprecise version
of foundational results such as the Gleason and Kochen-Specker theorems.
\end{abstract}

\tableofcontents

\chapter{Documentation}

This pdf doubles as documentation for the iuphd.cls file, and as a sample for the output.
The iuphd class is based on the standard \LaTeX \ report class, and as such you may assume that it follows
the default formatting of the report class unless otherwise specified below.  The
changes to the report class include the following:

\begin{itemize}
 \item changes to the default settings for margins, font size (11pt), spacing (double), and pagination,
 \item the quote, quotation, and verse environments have been changed to switch temporarily to single spacing, and
       remove the extra space at the top and bottom,
 \item the enumerate and itemize environments have been changed to remove extra space between the items and at the
       top and bottom,
 \item a line has been added to prevent the double spacing from affecting arrays,
 \item both numbered and unnumbered chapter headings are now centered by default,
 \item heading names such as CHAPTER, and BIBLIOGRAPHY are now all caps by default regardless of how they are typed,
 \emph{(This is accomplished with the help of textcase.sty.  Not all latex distributions come with this by default.)}
 \item changes to the definitions of \textbackslash maketitle and the `abstract' environment,
 \item the addition of commands for creating the acceptance page and copyright page,
 \item the addition of the `acknowledgments', `preface', and `dedication' environments,
 \item the addition of macros for specifying the department or school name, the names of
 the committee members, the defense date, and the copyright year,
 \item document class options for changing ``Department'' to ``School'' on the title page, and for showing
 the abstract, preface, acknowledgments, or dedication in the table of contents.
\end{itemize}

The appropriate commands for using these features are already typed into the dissertation.tex sample file,
with the optional items (dedication, preface, and copyright) commented out.  Simply remove the \% from the front
of these optional commands if you wish to use them.  The headings for the acknowledgments, preface, and dedication have
been defined as unnumbered chapter headings, which means they will not show up in the table of contents by default.
See the instructions below for adding them to the table of contents if desired.

\section{Current Standards at IU}
The current version of iuphd.cls complies with the standards of formatting a dissertation at IU as of 2013.  Below is a discussion of
some of the main points.

The current pagination standard is for the page number to appear centered at the top or bottom of the
page, and to use lowercase roman numerals for the material preceding the first chapter with the title page unnumbered but counting as
page one. The only part of this that is not standard in the \LaTeX \ report class is the use of roman numerals, which was accomplished
by changing the definitions to \textbackslash maketitle and \textbackslash tableofcontents.

The current standard for chapter headings seems to be that they are centered and in all caps.  The iuphd.cls enforces these
conventions automatically, however it is possible to override these changes with titlesec and titletoc. Because the textcase package is used,
the capitalization is not affected in math mode, so for instance
\medskip

\textbackslash section\{\$p\$-adic numbers\}
\medskip

\noindent will give the output ``$p$-ADIC NUMBERS'' in both the section heading and in the table of contents
(unless the value of \textbackslash tocdepth is changed to prevent sections from being listed in the table of contents).
Since all caps takes up more space horizontally than mixed upper and lowecase text, the size of the chapter headings
has been reduced to \textbackslash LARGE from \textbackslash huge.  However, no changes have been made to the size of any of the
other headings in the report class.

Most dissertations are one-sided, with an additional half inch added to the left hand margin on each page for the purpose of binding.
Although this is the standard, they do allow for the possibility of two-sided printing, which would cause the side with the extra margin
to alternate for binding.  This is automatically enforced simply by using the twoside option (a feature already built into the
standard report class). This is done as follows:
\medskip

\textbackslash documentclass[twoside]\{iuphd\}
\medskip

There does not seem to be a standard about whether the abstract, preface, acknowledgments, and dedication are listed in the table of
contents (TOC) or not.  As such, an option has been provided for including or excluding each one individually.  The syntax for these options
follows the pattern `showX' if X is to be listed in the TOC, and 'hideX' if X is to be omitted from the TOC, where X can be any of the four
items listed above.  In all cases, the default is for these \emph{not} to have entries in the table of contents.  So for example, to include
both the abstract and the preface in the table of contents, but not the dedication or acknowledgments, you would use the code:
\medskip

\textbackslash documentclass[showabstract,showpreface]\{iuphd\}
\medskip

\noindent To include a line for the bibliography in the table of contents, the command
\medskip

\textbackslash addcontentsline\{toc\}\{chapter\}\{BIBLIOGRAPHY\}
\medskip

\noindent should be placed directly above the \textbackslash bibliography and \textbackslash bibliographystyle commands.  It may be necessary
to use \textbackslash clearpage above this command to get the page number displayed in the contents to be correct.  However, if you are using
\textbackslash include for each chapter, then \textbackslash clearpage has technically already been used in this case, and so it is not
necessary to use it again. Similar commands will work for including any unnumbered heading in the table of contents, with careful attention
to the effects of \textbackslash clearpage.

Various commands have been defined for the data that will show up on the title page, acceptance page, and copyright page.  These include:
\begin{itemize}
 \item \textbackslash title - for the title of the dissertation
 \item \textbackslash author - for the name of the author of the dissertation
 \item \textbackslash date - for the \emph{completion date} of the dissertation
 \item \textbackslash department - this is optional.  The default value is ``Mathematics,'' but if you are in a different department, this
        command can be used to change its value to the name of whatever department (or school) you are in.  
 \item \textbackslash committeechair, \textbackslash readertwo, \textbackslash readerthree, \textbackslash readerfour - for the names of the
                      dissertation committee members.
 \item \textbackslash defensedate - for the \emph{defense date} of the dissertation (which may not match the completion date).
 \item \textbackslash cryear - for the copyright year.
\end{itemize}

Additionally the commands \textbackslash acceptancepage and \textbackslash copyrightpage have been defined to generate the acceptance page and copyright page
respectively.  The copyright page is not required, so it has been commented out of my dissertation.tex sample file.  Two document class options have been provided
for switching the text on the title page between ``Department'' and ``School.''  These options are given the obvious names `department' and `school'. The
default option is `department', so it only needs to be changed if you are in a school rather than a department.

A CV is also required following the bibliography, however, to maintain maximum flexibility I have not provided any special tools for
formatting a CV.  Various packages are available that do this, although I prefer custom formatting. As such, I have provided only a
line for including the CV after the bibliography, which is currently commented out.

\section{Suggestions for additional formatting}

For longer documents, it is often easier to manage if it is broken into multiple files.  This can be done by using the
\textbackslash include\{filename\} or \textbackslash input\{filename\} commands.  You can think of
\textbackslash input\{filename\} as an alias for all of the text contained in filename.tex: the text contained in that file,
simply gets inserted at the exact spot where the command \textbackslash input is used.  This means that filename.tex should not have
\textbackslash begin\{document\}, \textbackslash end\{document\}, or a preamble, as illustrated by iuphd-doc.tex in this example.
On the other hand, \textbackslash include\{filename\} has the additional feature of adding \textbackslash clearpage both before and after
the text being inserted.  This has important consequences for formatting.  Say that you have a file for each chapter,
and your first chapter is the introduction.  If you do something like this:
\medskip

\textbackslash chapter\{Introduction\}

\textbackslash include\{chap1\}
\medskip

\noindent then the title is formatted on its own page, and the text in chap1.tex will start on a new page; it produces
the same result as
\medskip

\textbackslash chapter\{Introduction\}

\textbackslash clearpage

\textbackslash input\{chap1\}
\medskip

\noindent To avoid having the title put on a separate page there are two options: use input instead of include
(without \textbackslash clearpage), or use the command \textbackslash chapter\{INTRODUCTION\} inside the file chap1.tex.
I favor the second option as illustrated in this example with iuphd-doc.tex.

Many \LaTeX \ packages are already available for additional customization of the formatting which is otherwise built in.
Some of these are listed below.  To use them, simply declare them with \textbackslash usepackage in the preamble
and refer to the documentation files for them, which are easily available online.

\begin{itemize}
 \item titlesec - for custom formatting of part, chapter, and section headings.
 \item titletoc or tocloft - for custom formatting of the table of contents.
 \item nomencl - for lists of notation (a tweak of makeindex).
 \item prettyref - for custom formatting of references (e.g. referencing equations as (1), (2), etc., instead of the default
       1, 2, etc.).  The ams packages also have a similar feature but it is much more limited.
 \item natbib, custom-bib, or biblatex - offer additional flexibility for bibliography formatting beyond the standard used by
       the mathematics community.
 \item setspace - contains features for resetting the global line spacing, as well as environments that can be used to change
       spacing locally, which are compatible with the quote, quotation, and verse environments.
 \item quoting - provides a more flexible quotation environment, with options for adjusting the margins manually.  This package
     should be used with setspace, to ensure single spacing of the quotations.
 \item enumitem or paralist - provide several more flexible list environments.
 \item titling - for custom changes to maketitle.
 \item geometry - for customizing the page layout.
\end{itemize}

%add features for block quotes, underlined titles, and different departments.

The last two should not be necessary unless the standards at IU change.  However, if that happens, it would really be better to
rewrite the iuphd document class.  Additional packages and other useful information can be found in \cite{mittelbach2004latex}.

>>>>

\documentclass[twoside]{iuphd}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[ backend=biber]{biblatex}
\addbibresource{prop.bib}
\ExecuteBibliographyOptions{sorting=none,maxbibnames=10,doi=false,isbn=false,url=false}
\DeclareUnicodeCharacter{0229}{\c{e}}
\DeclareUnicodeCharacter{00A0}{~}
\usepackage{mmap}
\usepackage{comment}
%\usepackage[basic]{complexity}
\newcommand{\lang}[1]{{\ensuremath{\mathsf{#1}}}}
\newcommand{\newlang}[2]{\newcommand{#1}{\lang{#2}}}
\newlang{\usat}{UNIQUE-SAT}
\usepackage[all]{xy}
\usepackage{framed}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[braket,qm]{qcircuit}
\usepackage{mathrsfs}
\usepackage{amssymb}

\newcommand{\says}[3]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{#1}{#2 says: #3}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\says{purple}{Yu-Tsung}{#1}}
\newcommand{\muB}{\ensuremath{\mu^{B}}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}

\begin{document}
%data for Title Page
\title{Discrete Quantum Theories and Computing}
\author{Yu-Tsung Tai}
\date{\today}
\maketitle[Department of Mathematics][Department of Computer Science]

%data for Acceptance Page
\defensedate{Defense Date}
\acceptancepage{Prof. Amr A. Sabry}{Prof. Dylan Paul Thurston}{Prof. Gerardo Ortiz}{Prof. Andrew J. Hanson}[Prof. Shouhong Wang]

%data for Copyright Page
\cryear{2018}
\copyrightpage

@@

\begin{abstract}
Our primary research interest is to build a quantum computing model
characterizing realistic quantum computers. While most of the quantum
computing models based on uncomputable numbers, that is, the continuum
of real numbers, most of the classical computers in our daily life
are digital instead of analog computers. This highlight the necessity
to investigate discrete models for quantum theory and computing. Specifically,
we start from replacing the continuum of complex numbers by the discrete
finite fields. Although we have fruitful results on their geometric
implications and computing powers, their probability models are still
not completely satisfactory. To address this issue, we further exploited
quantum interval-valued probability, and proved an imprecise version
of foundational results such as the Gleason and Kochen-Specker theorems. 
\end{abstract}
\tableofcontents{}

\chapter{Introduction}

Before diving into quantum computing, it is instructive to review
two kinds of classical computers: digital and analog. Although discrete
digital computers could be modeled by various computational models,
including Turing machines~\cite{Turing_1937}, these models are essentially
the same by the (strong) Church-Turing thesis \cite{Church_1936,Turing_1937,Kleene_1943}.
Moreover, any Turing-computable function could be carried out by a
digital computer as long as it has enough resources~\cite{Piccinini2015}.
In contrast, there are various idealized real computation models \cite{Siegelmann1998,Ziegler2007,weihrauch2012computable,blum2012complexity}.
Although these models are usually more powerful than discrete Turing
machines, their computational powers might not be realized by realistic
analog chips and computers, and the realized analog computational
devices are mostly chips embedded in our digital computers to communicate
with non-computational analog devices~\cite{Camenzind2005}. Currently
most of the quantum computing models based on the continuum of real
numbers. Similar to the development classical computers, the computational
power of realistic quantum computers might not be the same as predicted
by current quantum computing models. When developing realistic quantum
computers or estimating the computational powers of realistic quantum
computers, it might also be helpful to study discrete quantum computing
models.

Despite the history of computation, when replacing the continuum of
complex and real numbers by the discrete finite fields and finite
number of intervals, we show how a number of subtle properties of
quantum computing can be teased apart, step by step, as we explore
the implications of \emph{discrete quantum theories} in a systematic
fashion. In particular, on different discrete quantum theories, we
will present how geometrical structure of states and quantum probability
could be, and how these mathematical structures affect the power of
discrete quantum computing.

\chapter{Conventional Quantum Theory}

\section{Geometrical Structure of States}

An $n$-qubit pure state in the conventional quantum theory (CQT)
is represented as a vector in the $2^{n}$-dimensional Hilbert space.
If we eliminate all symmetries, an irreducible state is actually a
point in the projective Hilbert space \cite{MosseriDandoloff2001,Jaeger2007}
also known as the complex projective space~$\mathbb{CP}^{2^{n}-1}$
\cite{Hatcher2001,Bengtsson2007}. These irreducible quantum states
can also be classified as product states and entangled states, and
the latter one plays the essential role for pure-state algorithms
\cite{Mermin2007,Jaeger2007}.

\section{Quantum Probability}

Given a pure state~$\ket{\phi}$, when we measure an observable represented
by an Hermitian matrix~$\mathbf{O}$, the measurement result is one
of the eigenvalues of $\mathbf{O}$. The probability of getting a
particular eigenvalue~$\lambda$ is $\melem{\phi}{P}{\phi}$, where
$P$ is the projection operator onto the eigenspace of $\lambda$.
This rule of computing the probability is called the Born rule \cite{Born1983,Mermin2007,Jaeger2007},
which is used when we want to extract information from a quantum computer.
For any mixed state~$\rho$, the generalized Born rule induces a
conventional quantum probability measure $\muB_{\rho}:\events\rightarrow[0,1]$,
where $\events$ is the set of all projection operators on a given
Hilbert space. Conversely, any quantum probability measure $\mu:\events\rightarrow[0,1]$
can be induced from a mixed state $\rho$ in the Hilbert space of
dimension $d\ge3$ according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
In another word, this state $\rho$ is the unique state consistent
with any given quantum probability measure.

\chapter{Quantum Theories and Computing over Finite Fields}

\section{Modal Quantum Theory and Computing}

Our first discrete model replaces the complex numbers by a unrestricted
finite field~$\mathbb{F}_{q}$ \cite{Schumacher2012-SCHMQT,DQT2014,SchumacherWestmoreland2010}.
In this model, an $n$-qubit state is a non-zero vector in $\mathbb{F}_{q}^{2^{n}}$.
When we measure a given state~$\ket{\phi}$, an observable is replaced
by a basis~$\mathcal{B}$ of $\mathbb{F}_{q}^{2^{n}}$. Since $\ket{\phi}$
can be represented by the summation across a unique subset~$\mathcal{S}$
of the basis~$\mathcal{B}$, whether it is possible or impossible
to measure a basis vector~$\ket{i}$ depends on whether $\ket{i}$
is in $\mathcal{S}$ or not. Because this model only predicts whether
a result is possible or impossible, it is called the modal quantum
theory. Its computational model, called the modal quantum computing,
is far from conventional quantum computing. Although modal quantum
computing can express simple algorithms such as quantum teleportation
\cite{BennettBrassardEtAl1993,peres1995quantum,Mermin2007,Jaeger2007},
it is so weak that it cannot express Deutsch's algorithm \cite{Deutsch1985,Mermin2007}.
This quantum theory is, however, also so powerful that it can be used
to solve an unstructured database search of size $N$ using $O(\log(N))$
steps, which outperforms the known asymptotic bound $O(\sqrt{N})$
in conventional quantum computing \cite{Grover:1996:FQM:237814.237866,BennettBernsteinBrassardVazirani1997,Mermin2007,Jaeger2007}.

\section{Discrete Quantum Theory and Computing (I)}

Our second model called discrete quantum theory (I) considers only
finite fields of order $p^{2}$, with the prime $p$ of the form $4\ell+3$
($\ell$ a non-negative integer) \cite{geometry2013,DQT2014}. In
this model, an irreducible state is a vector in $\mathbb{F}_{p^{2}}^{2^{n}}$,
and can be reduced to a point in the discrete complex projective space~$\mathbb{DCP}^{2^{n}-1}$.
Among these irreducible states, the product states and entangled states
can not only be identified as in CQT but also be counted for different
$p$. Although the state space is a natural discrete analog to CQT,
this model still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for the unstructured
database search problem for certain database sizes.

\section{Discrete Quantum Theory and Computing (II)}

Our third model, discrete quantum theory (II)~\cite{DQT2014}, restricts
states in some local region of $\mathbb{F}_{p^{2}}^{2^{n}}$. Within
a local region, a notion of inner product and probability could be
recovered. Its discrete quantum computing can be further applied to
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.

\section{Toward Discrete Quantum Probability}

When people tried to define quantum probability over finite fields,
people tended to treat the original Born rule as an axiom, and tried
to modify it to get a discrete Born rule \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644,DQT2014,Ellerman2016a}.
However, any modified Born rule could hardly work on the whole vector
space, since there is no inner product on the whole vector space over
finite fields. Instead of treating the Born rule as an axiom, the
Born rule can actually be deduced from a set of abstract definitions
and axioms according to Gleason's theorem. Although we might hope
to deduce a discrete Born rule directly from a similar set of definitions
and axioms, no discrete Born rule satisfies certain properties motivated
by Gleason's theorem with infinitely precise real-number probability~\cite{Gardiner2014}.
Since the state spaces are now discrete and finite, this suggests
us to consider a discrete Born rule mapping to finite number of intervals
called interval-valued probability~\cite{JamisonLodwick2004,THOS2017}.
To adopting the idea of interval-valued probability step-by-step,
before attempting to study quantum interval-valued probability over
finite fields, we will first review the classical interval-valued
probability, and extend it with the conventional quantum theory.

\chapter{Quantum Interval-Valued Probability}

\section{Classical Interval-Valued Probability}

In the classical setting, there are several proposals for ``imprecise
probabilities'' \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
Although these proposals differ in some details, they all share the
fact that the probability $\bar{\mu}(E)$ of an event~$E$ is generalized
from a single \emph{real number} to an \emph{interval}~$[l,r]$,
where~$l$ intuitively corresponds to the strength of evidence for
the event~$E$ and~$1-r$ corresponds to the strength of evidence
against the same event. Given a sample space~$\Omega$ and a set
of intervals~$\mathscr{I}$, similar to a classical probability measure
$\mu:2^{\Omega}\rightarrow\left[0,1\right]$, a classical interval-valued
probability measure (IVPM) $\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
needs to satisfy some coherent axioms. By satisfying the convexity
axiom \cite{Shapley1971,GilboaSchmeidler1994,Marinacci1999,Grabisch2016},
Shapley proved that there is always a classical probability measure
consistent with the classical IVPM~$\bar{\mu}$ \cite{Shapley1971,GilboaSchmeidler1994,Grabisch2016}.
Given any random variable, its expectation value with respect to classical
probability measures consistent with~$\bar{\mu}$ is consistent with
its Choquet integral \cite{Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with respect to $\bar{\mu}$ \cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}.

\section{Quantum Interval-Valued Probability}
